The YUIMA Project: a Computational Framework for Simulation and Inference of Stochastic Differential Equations

### 2011-09-29 pkg/yuimadocs/inst/doc/JSS/article.Rnw
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@191 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

Adaptive Bayes estimation is developed in \pkg{yuima} via the method $\tt adaBayes$.
Consider again the model \eqref{eq:model1} with the same values for the parameters.
In order to perform Bayesian estimation, we need to prepare the prior densities for the parameters.
For simplicity we use uniform distributions in $[0,1]$
<<echo=TRUE>>=
prior <- list(theta2=list(measure.type="code",df="dunif(theta2,0,1)"), theta1=list(measure.type="code",df="dunif(theta1,0,1)"))
@
Then we call $\tt adaBayes$, on the same sample data we used for the \code{qmle} function, as follows
<<eval=TRUE,echo=TRUE, results=hide>>=
bayes1 <- adaBayes(yuima, start=param.init, prior=prior)
@
and we can compare the adaptive Bayes estimates  with the QMLE estimates
<<eval=TRUE,echo=TRUE>>=
coef(bayes1)
coef(mle1)
@
%The argument \code{method="nomcmc"} in \code{adaBayes} performs numerical
%integration, otherwise MCMC method is used.

>>>>

Adaptive Bayes estimation is developed in \pkg{yuima} via the method $\tt adaBayes$.
Consider again the model \eqref{eq:model1} with the same values for the parameters.
In order to perform Bayesian estimation, we need to prepare the prior densities for the parameters.
For simplicity we use uniform distributions in $[0,1]$
<<echo=TRUE>>=
prior <- list(theta2=list(measure.type="code",df="dunif(theta2,0,1)"), theta1=list(measure.type="code",df="dunif(theta1,0,1)"))
@
Then we call $\tt adaBayes$, on the same sample data we used for the \code{qmle} function, as follows
<<eval=TRUE,echo=TRUE, results=hide>>=
bayes1 <- adaBayes(yuima, start=param.init, prior=prior)
@
and we can compare the adaptive Bayes estimates  with the QMLE estimates
<<eval=TRUE,echo=TRUE>>=
coef(bayes1)
coef(mle1)
@
%The argument \code{method="nomcmc"} in \code{adaBayes} performs numerical
%integration, otherwise MCMC method is used.
\subsubsection{Small sample size}
It is known from the theory that the estimation of the drift in a diffusion process strongly depend on the length of the observation interval $[0,T]$.
In our example above, we took $T=n^(1/3)$, with $n = \Sexpr{n}$, which is approximatively  \Sexpr{round(n^(1/3),2)}. Now we reduce the sample size to $n=500$ and the value of $T$ is then $T=\Sexpr{round(500^(1/3),2)}$.
We then apply both quasi-maximum likelihood and adaptive Bayes type estimators to these data
<<>>=
n <- 500
ysamp <- setSampling(Terminal = n^(1/3), n = n)
yuima <- setYuima(model = ymodel, sampling = ysamp)
set.seed(123)
yuima <- simulate(yuima, xinit = 1, true.parameter = list(theta1 = 0.2, theta2 = 0.3))
param.init <- list(theta2=0.5,theta1=0.5)
mle2 <- qmle(yuima, start =param.init , 
lower = list(theta1=0, theta2=0), 
upper = list(theta1=1, theta2=1))
bayes2 <- adaBayes(yuima, start=param.init, prior=prior)
@
and we look at the estimates
<<>>=
coef(bayes2)
coef(mle2)
@
Compared to the results above, we see that the parameters in the diffusion coefficients are estimated with good quality while for the parameters in the drift the quality of estimation deteriorates. The adaptive Bayes estimator seems to perform a little better though.
